\documentclass[../main.tex]{subfiles}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------


%\chapter*{Bibliography}
\addcontentsline{toc}{chapter}{\textcolor{ocre}{Bibliography}}

%\bibliographystyle{unsrt}

%\section*{Books and Films}
%\addcontentsline{toc}{section}{Books}
%\printbibliography[heading=bibempty,type=book]

%\section*{Articles}
%\addcontentsline{toc}{section}{Articles}
%\printbibliography[heading=bibempty,type=article]

%\section*{Online Resources}
%\addcontentsline{toc}{section}{Webpages}
%\printbibliography[heading=bibempty,type=article]

\begin{thebibliography}{9}

\bibitem{AR}
Angevine, Robert.
\textit{The Railroad and the State: War, Politics, and Technology in Nineteenth-Century America}
Stanford: Stanford University Press, 2004.

\bibitem{ALW}
Arps, Louisa Ward.
2007.
\enquote{Denver in Slices: A Historical Guide to the City.}
Colorado State.
\\\url{http://lib.colostate.edu/publicwiki/images/e/eb/Colorado\%26FrontRangebooks.pdf}

\bibitem{fate}
Balsera, Leonard, Brian Engard, Jereky Keller, Ryan Macklin, and Mike Olson.
\textit{Fate: Core System, Fourth Edition.}
Evil Hat Productions,
2013

\bibitem{QCR2}
Brownell, Blaine A.
1978.
Review of \textit{Review of the Queen City: A History of Denver}, by Lyle W. Dorsett.
\textit{The American Historical Review}
83 (4): 1099?1100. doi:10.2307/1867833.

\bibitem{DWW}
\textit{Dances with Wolves},
directed by Kevin Costner.
(2003, Santa Monica, CA: MGM Home Entertainment), DVD.

\bibitem{RDO}
Dunbar-Ortiz, Roxanne.
\textit{An Indigenous People's History of the United States (ReVisioning American History)}.
Beacon Press, 2015.

\bibitem{EoM}
\enquote{Encyclopedia of Mormonism?:: Encyclopedia of Mormonism.}
Accessed February 20, 2016. 
\\\url{http://contentdm.lib.byu.edu/cdm/compoundobject/collection/EoM/id/4391/show/4298.}


\bibitem{PE}
Goddard, Pliny Earle.
\textit{Jicarilla Apache Texts, 1911.}
From New York: Anthropological Papers of the American Museum of Natural History, Vol. VIII.
Accessed May 7th 2016.
\url{http://www.sacred-texts.com/nam/pla/jat/ }

\bibitem{butch}
\textit{Butch Cassidy and the Sundance Kid.}
Directed by George Roy Hill.
(2006, Beverly Hills, CA: 20th Century Fox Home Entertainment), Streaming Video.

\bibitem{hope}
\textit{The West: The Geography of Hope (1877-1887).}
Directed by Stephen Ives. 
(1996, Bethesda, MD: PBS.) Streaming Video.

\bibitem{sss}
Jackson, Mary Susan
 1978.
 Review of \textit{Review of the Queen City: A History of Denver}, by Lyle W. Dorsett.
\textit{The Journal of American History}
65 (2): 489?90. doi:10.2307/1894156.

bibitem{dead}
\textit{Dead Man}
Directed by Jim Jarmusch.
(1995, Santa Monica, CA: Miramax Films). DVD.

\bibitem{JM}
Jones, Mary Ellen
\textit{Daily Life on the Nineteenth Century American Frontier.}
Greenwood, 1998.

\bibitem{gbu}
\textit{The Good, the Bad, and the Ugly}
Directed by Sergio Leone
(1967, Italy: Produzioni Europee Associati) DVD

\bibitem{VE}
Lehdonvirta, Vili, and Edward Castronova.
\textit{Virtual Economies: Design and Analysis.}	
Cambridge, MA: The MIT Press, 2014.

\bibitem{CAI}
Liestman, Daniel. 1999.
\enquote{Horizontal Inter-ethnic Relations: Chinese and American Indians in the Nineteenth-century American West}. 
\textit{The Western Historical Quarterly}
30 (3). Utah State University: 327?49.
doi:10.2307/971376.

\bibitem{BM}
McCarthy, Cormac.
\textit{Blood Meridian: Or the Evening Redness in the West.}
New York: Vintage, 1992.

\bibitem{QCR4}
Morris, John R. 1978
Review of \textit{Review of The Queen City: A History of Denver}, by Lyle W. Dorsett.
\textit{The Western Historical Quarterly} 9 (3): 367?367.
doi:10.2307/966943.


\bibitem{OE}
\enquote{Order of Enoch.} Accessed February 20, 2016. 
\\\url{ttp://mormonthink.com/glossary/orderofenoch.htm.}

\bibitem{Sum}
Saunders, Greg. 2010.
\textit{Summerland Revised and Expanded Edition.}
Cubicle 7 Entertainment  Ltd.

\bibitem{RB}
Taillon, Paul Michel.
\enquote{'What We Want Is Good, Sober Men:' Masculinity, Respectability, and Temperance in the Railroad Brotherhoods, C. 1870-1910}
\textit{Journal of Social History}
36, no. 2 (2002): 319?38.

\bibitem{Sigam}
Turner, Frederick Jackson. \enquote{The Significance of the Frontier in American History.}
Henry Holt and Company,
1921.

\bibitem{MT}
Twain, Mark (Samuel Clemens).
\enquote{Roughing It}
August 18, 2006 [EBook \#3177]. 
\\\url{http://www.gutenberg.org/files/3177/3177-h/3177-h.htm.}

\bibitem{QC}
W, Dorsett Lyle, and McCarthy Michael.
1986.
\textit{The Queen City: A History of Denver.}
2nd edition. 
Boulder, Colo: Westwinds Press.

\bibitem{Stage}
\textit{Stagecoach}
Directed by John Ford
(Burbank, CA: Warner Home Video) Streaming Video.

\bibitem{CWR}
West, Elliot.
\enquote{The Civil War and Reconstruction in the American West.}
The Gilder Lehrman Institute of American History.
Accessed May 7th 2016.
\\\url{https://www.gilderlehrman.org/history-by-era/reconstruction/essays/civil-war-and-reconstruction-american-west}


\end{thebibliography}