%----------------------------------------------------------------------------------------
%	CHAPTER CONFLICT
%----------------------------------------------------------------------------------------
\documentclass[../main.tex]{subfiles}

\chapter{Conflict Themes}

	\lettrine[lines=2]{A}{merican} Sunset is driven by three main sources of conflict, sin and corruption, chaos and disorder, and evil and emptiness. These three sources of conflict are made manifest in the people and places in the West. That's not to say every conflict is simply one of these, many conflicts can be much larger and more complex. Furthermore the people of the West pushing the conflicts forward might themselves be unaware of where their motivations lie. That's not to say that each and every conflict is the result of some large scale abandonment of morals, but simple greed, rebellion or just a bad person can cause many of the simple conflicts that are found in our setting. Bandits may be driven by need or greed, but even those who become bandits out of desperation have probably suffered at the hands of sinful, chaotic, or evil circumstances. 

\hfill

\textbf{Sin and Corruption} in the Hearts of Men. The push into the frontier made room for new figures of influence to arise be they political, economic, or spiritual. The desire for power and influence is present in many figures that shaped the west, and in the course of your character's journey you will surely find yourselves opposed to them. G.M. Holt is a man of wealth and power, who is distributing his bribes amongst the pockets of others like himself, to gain political, social or further economic influence. His money buys politicians and businessmen and women, bringing his interests to bear over the interests of the community. But Silver Springs is a town founded on the idea of mineral prosperity, and greed is all but inseparable from the town's history. Holt is by no means unique by his avarice, nor is anyone else by their sin of choice. As a player, bear in mind that even the most noble of causes may be playing into the hands of the vilest of Silver Springs denizens. 

	Players playing games in settings heavily influenced by sin and corruption may find themselves surrounded by religious NPCs, a strong Enochite or christian presence, biblical references and symbolism, and even some strange biblical magic. Corruption in the game might be of an ethical nature, like political corruption, or of a moral nature like corruption of someone's character, causing a change in their behavior. In the latter case it could be anything from a large dose of cynicism, to something more sinister.

\hfill

\textbf{Chaos and Disorder} in the confusion of expansion. In the west, multitudes of different parties with different interests were thrown together, and out of that chaos arises an inherent conflict. Anyone can be sucked into this maelstrom at any time, and to unravel the twisted threads is not a simple task. Between Steele, who cannot leave the civil war behind and Tecumseh riding to this death, the bandits that can attack anywhere at anytime, and the lack of strong law in town, anything can happen at any time and the consequences can be incredibly unpredictable. The Civilized Five claim it to be the Trickster at work, taking advantage of the ignorance of the settlers, while the successful people of Silver Springs describe it as opportunity, opportunity to be the first of something, the best at something or to control the most of something. The lack of an appropriate system of checks and balances in the West makes even simple tasks seem to have repercussions far outside what a rational mind might expect.

	Players playing games in settings heavily influenced by chaos and disorder may find themselves surrounded by NPCs strongly related to the law, like sheriffs, vigilantes, and bandits, as well as  strong ambitious leaders trying to take control of the situation. Furthermore, a lot of native american mythology has to do with the Trickster, and players may find characters using superstitious explanations to account for odd, unpredictable events, including angry spirits, skinwalkers, or the Trickster himself. Often people profiteering from the disorder and disruption may be working behind the scenes to further entangle the players. These profiteers could be actual people, like politicians and business people, or they could be entities just wanting to watch the world burn.

\hfill

\textbf{Evil and Emptiness} in the vast reaches of the frontier So much of the west is unknown; towns are few and far between. In those gaps where civilization is stretched thin there is space for all nature of unsavory characters with no more motivation than their own twisted entertainment. People tell stories of cultists and heathens who lost their civilized minds in the frontier and live alone in the woods as animals, preying on travelers. Others of course form communities in the empty places and practice unsavory religious rituals and lifestyles, or so it is said. True or not there is something very similar about the vast nothingness that Silver Springs buffers against. Where once there were nomadic peoples and herds of bison roaming everywhere, now the chances of meeting these people or seeing majestic creatures is far lower. The unprepared often go missing, perhaps eaten by wolves, choking on poisonous berries, dying of thirst and starving. Still others are asked to confront something evil about themselves as they do anything in their power to survive. Such experiences are life changing and those who go out and come back from the emptiness, especially when they return alone, are often kept at a distance from society, by their choice or that of the community.
	
	Players playing games in settings heavily influenced by a sense of emptiness and evil may find themselves mistrusted by many NPCs and more frequently encounter people who have slipped through the cracks, such as small unsavory cults, the sociopathic or the criminally insane. In investigating the empty places they may come across revelations that are hard to bear, such as the humanity of Natives and the inhumanity of the American Army. In investigating the interactions between the people of the emptiness and the people still in touch with humanity they may also come across genuine mysteries with terrifying endings, such as cannibals, people who enjoy killing, or others things from beyond the rational mind...