\documentclass[../main.tex]{subfiles}

%----------------------------------------------------------------------------------------
%	CHAPTER MAGIC
%----------------------------------------------------------------------------------------



\chapter{Magic}

\lettrine[lines=2]{M}{agic} in American Sunset is a far cry from the wizardly powers of other settings. Subtle, pervasive, and motivated more by the whims of the world than the players, you should expect to find it wherever stories are told. Belief guides most otherworldly forces on the frontier, giving rise to trickster gods as readily as fallen angels, and to more terrestrial powers in the hands of players. While you shouldn't believe everything you're told, know that there's a little bit of truth in every story. In keeping with this, the magic of the West is split into two categories, Reputation and Mystery.


\section{Reputation}
	Reputation is the magic of the known and the magic most readily available to players. A representation of what you're believed to be by the people in American Sunset, your reputation is magic unto itself. In a land where stories travel faster than storytellers, the tales told of your accomplishments and their reality are divided by a thin line. While the magic of Reputation lies in the hands of the players, it is a subtle and specific kind of power. A player's past accomplishments give them the ability to influence the odds in their favor, shaping their reality into one that conforms to their reputation. Under the right circumstances, a player can make the impossible possible, because otherwise there'd be no stories to tell. These impossible feats are closely tied to their performer;  Expect onlookers to be impressed, even awestruck, but never surprised. If the Fastest Gun in the West kills seven men in the blink of an eye, then he's done the impossible, to be certain- but when it comes to shooting, nothing's impossible for the Fastest Gun in the West. Why else would he carry the name?
	
Take note of the stories told about your antagonists as you encounter them. The more often you hear them repeated, the more likely they're indicative of their abilities. 

\section{Mystery}
	 In contrast to Reputation, which is the magic of things that are known, Mystery is the magic of everything that is unknown. As colonization of the West proceeds, more settlers will find themselves pushing against forces they cannot understand. Mystery cannot be mapped by cartographers, or catalogued by the instruments of modern science, but they can still be understood through the oral traditions of those who live alongside it. If you hear rumors that a local mountain is haunted, and these stories are supported by locals, or by tales told by the natives who live in the area, think twice before moving forward.
	 
Strange creatures and stranger forces are at play in the unmapped lands. Chupacabras still stalk through the desert flats to the South. The hoots and gibbering of hirsute wildmen of impossible stature still echo through the forests to the North. On the plains to the East, at the right time of year, on the right kind of night, the glowing bones of ancient creatures come alive, swimming through waves in a sea of grass. Still more dangerous things walk like men between the scattered villages and homesteads on the frontier, ancient and cunning and hungry.

And you might be one of the last to see these things firsthand. In the world of American Sunset, the frontier is just years away from closing. Mystery, while still a power in the wilds, is disappearing as quickly as the wilds themselves. Once mapped, the mystery in the land vanishes, leaving behind little more than landscapes where once ancient spirits hid beneath the hills. 
Those who've lived their lives in comfort behind the walls of cities and towns do not, and perhaps even cannot, understand what lies beyond the edges of the map. Do not expect your tales of haunted caves and impossible animals to be taken seriously within the city limits.