%----------------------------------------------------------------------------------------
%	INTRO
%----------------------------------------------------------------------------------------
\documentclass[../main.tex]{subfiles}

\vspace*{\fill}
%\begin{center}
%\begin{minipage}{.8\textwidth}
\lettrine[lines=2]{I}{n} 1890, following the eleventh U.S. Census, the United States government announced that the American frontier was closed. Three years later, historian Frederick Jackson Turner seized upon this report and declared that the end of an era had passed. The America that came after 1893 would be fundamentally different than the one that came before; the land was \enquote{settled,} lacking a wild and unpredictable edge.

The Western, the quintessential tale of the American frontier, is a romance told about the last days of the America that came before. Not quite real, a land where the people are larger than life, death is cheaper than a glass of whiskey, and the horizon is as endless as it is beautiful.

The reality, however, is sadly far less romantic.

The atrocities committed in the name of westward expansion are innumerable and horrific. The American Indian Wars led to the deaths of tens, if not hundreds of thousands and eroded the culture, religion, and way of life of Native American tribes, all in the name of bringing \enquote{civilization} to the frontier. The history of the American West is blood-stained and tragic.

The book you're reading is our endeavor to distill these complex histories of the West into a single, playable experience. No small amount of research went into creating this work, and even then, it was far from a simple task. What we've created isn't perfect, but it is the culmination of our efforts at giving you the opportunity to experience all of the West - for better, and for worse. In keeping with this, while the places, names, and organizations that follow are fictional, they are all inspired by true-life people and events.

American Sunset is a game about adventures, train robberies, outlaws, saloon shootouts, horseback-riding, mountain-scaling, river-fording, and a thousand more things, but it is also a game about loss. 

The loss of a way of life. The loss of a people. The loss of the America that came before, and the loss of the frontier on her border.

We hope that you have fun playing American Sunset, and we hope that you are able to experience the Western story you want to play, but as you do, we also hope that you keep these things in mind.

\hfill

Thank you, and enjoy.

%\end{minipage}
%\end{center}
\vfill % equivalent to \vspace{\fill}